### 2024-05-21 paper/setup.tex
Change focus to Implementation of an optimal branch-decomposition algorithm for planar graphs.

% General
	\usepackage{graphicx} % Required for inserting images
	\setlength{\parindent}{0pt}
	\setlength{\parskip}{8pt} % Vertical space between paragraphs
	\setlength{\baselineskip}{18pt} % Vertical space between lines
	\usepackage{amsthm}
	\usepackage{mdframed}
	\usepackage{hyperref} % Add the hyperref package for clickable links
	\newcommand\bigO{\mathrm{O}}
	\usepackage[T1]{fontenc}

% Define the theorem-like environments
	% Define a new theorem style with upright text
	\newtheoremstyle{plainupright} % Name of the style
	{10pt}                        % Space above
	{3pt}                        % Space below
	{}                           % Body font (default is \itshape which is italics)
	{}                           % Indent amount
	{\bfseries}                  % Theorem head font
	{.}                          % Punctuation after theorem head
	{ }                          % Space after theorem head
	{}                           % Theorem head spec (can be left empty, meaning 'normal')

	\theoremstyle{plainupright}

	\newtheorem{theorem}{Theorem}[section]  % Numbering within sections
	\newtheorem{definition}[theorem]{Definition}  % Numbering with theorems
	\newtheorem{corollary}[theorem]{Corollary}  % Numbering with theorems
	\newtheorem{claim}[theorem]{Claim}  % Numbering with theorems
	\newtheorem{lemma}[theorem]{Lemma}  % Numbering with theorems
	\newtheorem{axiom}[theorem]{Axiom}  % Numbering with theorems
	\newtheorem{conjecture}[theorem]{Conjecture}  % Numbering with theorems
	\newtheorem{problem}[theorem]{Problem}  % Numbering with theorems

% Enclose theorem-like environments within boxes
	\surroundwithmdframed[style=theoremstyle]{theorem}
	\surroundwithmdframed[style=theoremstyle]{definition}
	\surroundwithmdframed[style=theoremstyle]{corollary}
	\surroundwithmdframed[style=theoremstyle]{claim}
	\surroundwithmdframed[style=theoremstyle]{lemma}
	\surroundwithmdframed[style=theoremstyle]{axiom}
	\surroundwithmdframed[style=theoremstyle]{conjecture}
	\surroundwithmdframed[style=theoremstyle]{problem}

% Define a box style for the theorem-like environments
	\mdfdefinestyle{theoremstyle}{%
		linewidth=0pt,
		backgroundcolor=white,
		roundcorner=5pt,
		innertopmargin=6pt,
		innerbottommargin=0pt,
		innerrightmargin=32pt,
		innerleftmargin=16pt,
	}

% Set margins
	\usepackage{geometry}
	\geometry{
		left=3cm,
		right=3cm,
		top=2cm,
		bottom=2cm
	}

% Codeblock
	\usepackage{minted}
	\usepackage{xcolor}
	\definecolor{mintedBg}{rgb}{0.98,0.98,0.98}
	\setminted[py]{
		bgcolor=mintedBg,
		framesep=3mm,
		fontsize=\footnotesize,
		linenos,
		breaklines,
		breakanywhere,
	}
	\newcommand{\pyfile}[1]{%
		\textsf{%
			\detokenize{#1}%
		}%
		\par
		\inputminted{py}{../code/\detokenize{#1}}%
	}